---
title: "LAB 4"
format:
  html:
  embed-resources:true
---

Question 1: Download the Data

```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )

```

Question 2: Clean the Data

```{r}
met <- met[met$temp >= -16, , drop = FALSE]

met <- met[met$temp >= -16, , drop = FALSE]

met_avg <- met[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
), by=c("USAFID", "day")]
met <- na.omit(met)

met$region <- ifelse(met$lon < -98.0 & met$lat > 39.71, "NE",
ifelse(met$lon < -98.0 & met$lat <= 39.71, "SE",
ifelse(met$lon >= -98.0 & met$lat > 39.71, "NW", "SW")))
met_avg[, elev_cat := ifelse(elev > 252, "high", "low")]
met_avg[, elev_cat := ifelse(elev > 252, "high", "low")]

```

Question 3:

```{r}
library(ggplot2)
ggplot(met, aes(x = factor(1), y = wind.sp)) +
  geom_violin(fill = "blue", alpha = 0.5) +  # Wind speed
  geom_violin(aes(y = dew.point), fill = "green", alpha = 0.5) +  # Dew point
  facet_wrap(~region) +  # Facet by region
  labs(x = NULL, y = NULL) +  
  theme_minimal() +  
  theme(panel.grid = element_blank())  


```

-   Describe what you observe in the graph: I notice that the northeast region has the highest dew point. I can also see that the dew point and winds peed do not share any correlation.

Step 4:

```{r}
library(ggplot2)

# Remove rows with missing values in 'wind.sp' and 'dew.point'
met <- na.omit(met)

# Create a scatterplot with jittered points and linear regression lines by region
ggplot(met, aes(x = dew.point, y = wind.sp, color = region)) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.5) +  # Jittered points with color by region
  stat_smooth(method = "lm", se = FALSE, formula = y ~ x, aes(group = region)) +  # Linear regression lines by region
  labs(x = "Dew Point", y = "Wind Speed") +  # Axis labels
  theme_minimal()  # Minimal theme

```

-   Describe what you observe in the graph: I can see that the southwestern region has a higher wind speed compared to the other three regions. The dew points are relatively similar to one another, with the exception of a few outlires.

Step 5:

```{r}
library(ggplot2)

# Remove rows with missing values in 'elev' and 'region'
met <- na.omit(met)

# Create barplots of weather stations by elevation category
ggplot(met, aes(x = factor(elev), fill = region)) +
  geom_bar(position = "dodge") +  # Create bars with dodge positioning
  scale_fill_brewer(palette = "Set1") +  # Change colors using Brewer palette
  labs(x = "Elevation Category", y = "Number of Stations", fill = "Region") +  # Axis labels and legend title
  ggtitle("Weather Stations by Elevation Category and Region") +  # Title
  theme_minimal() +  # Minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

Step 6:

```{r}
library(ggplot2)


met <- na.omit(met)

# plot for dew point
ggplot(met, aes(x = region, y = dew.point)) +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar", width = 0.2) +
  stat_summary(fun.data = "mean_sdl", geom = "point", size = 3) +
  labs(x = "Region", y = "Dew Point") +
  ggtitle("Mean Dew Point by Region with Standard Deviation Error Bars") +
  theme_minimal()

#  plot for wind speed
ggplot(met, aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar", width = 0.2) +
  stat_summary(fun.data = "mean_sdl", geom = "point", size = 3) +
  labs(x = "Region", y = "Wind Speed") +
  ggtitle("Mean Wind Speed by Region with Standard Deviation Error Bars") +
  theme_minimal()

```

-   Describe the graph and what you observe: In the dew point graph, we can see that the southwestern region has the highest dew point levels. in the wind speed graph, we can see that all regions are relatively the similar and within close range of one another.

-   Dew point is... NE= 10, NW=16, SE= 12, SW= 22

-   Wind speed is...NW= 4, NW= 3.2, SE= 4.1, SW= 4.2

Step 7:

```{r}

library(leaflet)

met <- na.omit(met)

#color palette with custom colors
colors <- colorFactor(
  palette = c("blue", "green", "yellow", "red"),
  domain = met$region
)

# leaflet map
m <- leaflet() %>%
  addTiles() %>%
  setView(lng = -98.00, lat = 39.71, zoom = 4)  

#  markers for the top 10 places in relative humidity
top10 <- met[order(-met$rh), ][1:10, ]
m <- m %>%
  addMarkers(
    data = top10,
    lat = ~lat,
    lng = ~lon,
    label = ~paste0("Place: ", rownames(top10), "<br>Relative Humidity: ", round(rh, 2), "%"),
    clusterOptions = markerClusterOptions()
  )

# legend values
legend_values <- levels(colors)

# Add a legend
m <- m %>%
  addLegend(
    "bottomright",
    pal = colors,
    values = legend_values,
    title = "Region",
    opacity = 1
  )

# Display the map
m


```

-   Describe the trend in RH across the US: The northwestern and northeastern regions have the highest relative humidity in the US.

Step 8:

```{r}
install.packages("patchwork")
library(ggplot2)
library(patchwork)
# Create a scatterplot for Plot 1

plot1 <- ggplot(data = data.frame(x = rnorm(100), y = rnorm(100)), aes(x = x, y = y)) +
  geom_point() +
  labs(title = "Scatterplot 1")

# Create a scatterplot for Plot 2
plot2 <- ggplot(data = data.frame(x = rnorm(100), y = rnorm(100)), aes(x = x, y = y)) +
  geom_point() +
  labs(title = "Scatterplot 2")


```
